### 2014-11-29 paper/introduction.tex

\section{Introduction}
\label{sec:intro}

Modeling a predator's food selection was given much attention for a short period of time; see \citet{Strauss:1979} and the references within.  Many, if not all, of the indices developed, though intuitive, focused on a snapshot in time and on only one prey species, and rarely obtained practical asymptotic properties.  This restricts the interested practitioner to the most computationally manageable of techniques, and has completely ignored simultaneous testing across an array of both species and time.

The model presented here maintains tractable asymptotic properties while being general enough to take into account multiple species and time points.  By modeling both time and any number of prey species, we are able to see a more detailed analysis of the predator's eating preferences.  The simplicity of the model allows us to consider predators for which exact tallies of the number of each prey species eaten within any given time period is not observed.  Instead, we rely on the researcher being able to DNA sequence the contents of the predator's gut and make a simple binary conclusion: this predator ate some of that prey species during this time period, or did not.  Despite the data's lost information, under certain situations our model is able to accurately estimate the parameters of interest.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "main"
%%% End: