### 2015-01-13 paper/acknowledgments.tex
last of bio's comments

\section{Acknowledgements}
\label{sec:acknowledge}

The information reported in this paper (No.\ $15-08-008$) is part of a project of the Kentucky Agricultural Experiment Station and is published with the approval of the Director.  Support for this research was provided by the University of Kentucky Agricultural Experiment Station State Project KY008055 and the National Science Foundation Graduate Research Fellowship Program.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "main"
%%% End: