### 2015-12-10 paper/EEST/First_Submission/Biographies/EEST-D-15-00056_biographies.tex
Added author biographies for EEST.

\documentclass[12pt]{article}

\setlength{\parindent}{0pt}
\setlength{\parskip}{12pt}

\begin{document}
\begin{center}
  \begin{Large}
    \textbf{Author Biographies for EEST-15-00056:}\\
    \textsc{Formal Modelling of Predator Preferences using Molecular Gut-Content
      Analysis}
  \end{Large}
\end{center}

\textbf{Simon Bonner} is an Assistant Professor jointly appointed between the Department of Statistical and Actuarial Sciences and the Department of Biology at the University of Western Ontario, London, Ontario, Canada (sbonner6@uwo.ca). His research focuses primarily on developing statistical methods for the analysis of ecological data with particular emphasis on Bayesian methods for analyzing data from mark-recapture studies of wild animals. This work was completed while Simon was an Assistant Professor in the Department of Statistics at the University of Kentucky.

\textbf{James Harwood} is an Associate Professor of Insect Ecology in the Department of Entomology at the University of Kentucky, Lexington, Kentucky, USA (james.harwood@uky.edu). His research seeks to understand mechanisms of foraging by generalist predators and identify their role in biological control through the integration of molecular techniques, behavioral studies and field experiments.

\end{document}