### 2016-01-25 paper/EEST/Second_Submission/Response/EEST-D-15-0056R1_response.tex
plots to grey-scale

\documentclass[12pt]{article}

\usepackage[margin=2.0cm]{geometry}

\begin{document}
\begin{center}
  \begin{Large}
    \textbf{Response to Reviews of EEST-15-00056R1:}\\
    \textsc{Formal Modelling of Predator Preferences using Molecular Gut-Content
      Analysis}
  \end{Large}
\end{center}

\section{Reviewer \#1:}
\begin{enumerate}
\item  \textit{The authors responded everything I suggested at the first reviewing. I have no additional comment and recommendation, except one typing error; P4 L33 (L137) Yjst to Yist}

Fixed.

\end{enumerate}
\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End: